---
title: "hw2_fintech"
author: "Ruizhou Peng"
date: "`r Sys.Date()`"
output: html_document
---


To begin, install the 'quantmod' package and load stock price data from Yahoo Finance.
```{r eval=FALSE, include=FALSE}
install.packages('quantmod')
```

```{r}
library(quantmod)
library(tidyverse)
getSymbols('AAPL', src = 'yahoo')

df_AAPL <- data.frame(Date=as.character(index(AAPL)), coredata(AAPL))
```

## 1.Data Wrangling

q1: inspect the dataset with **head**, and check if there is any NA value in each column
```{r}
df_AAPL %>% head(5)
df_AAPL %>% summarise(across(everything(),
                             ~sum(is.na(.))))
```

q2: load stock price of MicroSoft using ticker "MS", transfer to data frame and combine the stock price of MS and AAPL into one data frame df_all. Select Date, Open and Close price of two stocks.

```{r}
# first load MicroSoft stock
# coredata--->extract the observation without time series
getSymbols('MS', src= 'yahoo')

df_MS <- data.frame(Date=as.character(index(MS)), coredata(MS))
```

```{r}
# use join to combine the stock price of MS,AAPL
# into one data frame

df_all <- df_AAPL%>%select(Date, AAPL.Open, AAPL.Close) %>%
  inner_join(df_MS%>%select(Date, MS.Open, MS.Close))

```

check
```{r}
df_all %>% head(5)
```

## 2. Data Wrangling and Visualization

q3: caculate the intraday return into \{stockname\}.R

```{r}
df_all <- df_all %>% mutate(AAPL.R= (AAPL.Close-AAPL.Open)/AAPL.Open)


df_all <- df_all %>% mutate(MS.R= (MS.Close-MS.Open)/MS.Open)
```

q4: plot the density of AAPL.R using **ggplot** and **geom_density**
```{r}
# use 'fill' arg to control filling color
df_all%>%ggplot(aes(AAPL.R))+geom_density(fill='blue')+xlab('AAPL.Return')+ylab('Density of Return')
```

q5: we need to extract the information of stock performance by year. First split column **Date** into "Year", "Month" and "Day" and create a new frame called **df_ymd** transformed by **df_all**
```{r}
df_ymd <- df_all%>%
  separate(Date, c("Year", "Month", "Day"), sep='-')
```
check 
```{r}
df_ymd %>% head(5)
```
q6: create a boxplot of MS.R using **ggplot** and **geom_boxplot**. To distinguish returns across different years, utilize the **aes()**
```{r}
df_ymd %>% 
  ggplot(aes(x=MS.R, y=Year))+geom_boxplot()
```
From this boxplot, we can see that the inner quartile range is shrinking across the year while the median of the returns stay steady nearly 0. The abnormal movements in the stock may lie in 2008 since it has a wide IQR and there are a lot of outliers, this phenomenon may reflect that the stock is going through some changes and is fluctuating.

## 3. Data Wrangling, Tidying and Visualization

q7: create a frame **df_sum** for the stocks' average return over months during 2000-2020

```{r}
df_sum <- df_ymd%>%
  filter(Year>=2000 & Year<= 2020)%>%
  select(Month, AAPL.R, MS.R)%>%
  group_by(Month)%>%
  summarise(AAPL.Mn=mean(AAPL.R, na.rm=TRUE),
            MS.Mn = mean(MS.R, na.rm=TRUE))
```

check
```{r}
df_sum %>% head(5)
```

q8: tidy df_sum since the columns, and generate bar-chart for AAPL.Mn and Ms.Mn over months
[non-tidy since columns represent a single variable]

```{r}
# utilize pivot_longer to transform columns to a single column: Stock
long_df <- df_sum %>%
  pivot_longer(!Month ,names_to = 'Stock', values_to = 'Returns')

```
check

```{r}
long_df %>% head(10)
```
then use **ggplot** to show bar-chart over different months
```{r}
# ggplot:
#   fill---> filling the color according to the specified column
#     using different values in the column

# position_dodge:
#   adjusting the horizontal position,need to be specified in the
#   global or geom* layer 
#   usage: position= position_dodge(width=?)

long_df %>% 
  ggplot(aes(x=Month, y=Returns, fill=Stock))+
  geom_col(position = position_dodge(width = 0.4), width=0.4)
```